Making (Secure) FETCH Happen: \\ \vspace{2 mm} {\large End-to-End Metadata Security for IMAP}

### 2015-04-29 paper/paper.tex
changes from paper readthrough

\begin{abstract}
Although communication metadata is known to be intrusive, most end-to-end encryption solutions focus only on hiding the contents of messages. Yet who a user is talking to and when is information as sensitive as the contents of the message. While end-to-end metadata security solutions exist for message transport, existing solutions lack the multi-device synchronization that IMAP provides. Without providing full feature parity with legacy email technologies, a metadata-secure protocol cannot hope to achieve ubiquity. To that end, we present \projectnospace, a client-side extension of the IMAP protocol held to rigorous standards of feature parity without sacrificing on security. With \projectnospace, existing IMAP clients can be invisibly upgraded to provide metadata security for email storage and synchronization.
\end{abstract}

@@

Today, most email is hosted by cloud providers. Unfortunately for users, this means that information about who they're talking to, when they're talking to them, and what they're saying is readily available to any interested company, government, or advertiser. Given that a modern threat model must account for the mail provider as an adversary, then only an end-to-end (E2E) solution will provide adequate security. Furthermore, even encrypting the contents of the email will protect only a small portion of the full breadth of sensitive information: the metadata of who is communicating and when is equally important to protect. Fortunately, recent projects have emerged to add protections for metadata into the communication security toolkit. Yet these projects lack a key feature to achieve full feature parity with insecure messaging solutions: support for multi-client synchronization and storage across all of a user's devices.

@@

@@

To send an email, the sender's machine sends a request to an SMTP server acting on its behalf, which relays it through other SMTP servers until it reaches the recipient's destination server. At that point, the recipient's machine retrieves the email from the server using either POP or IMAP.

\subsubsection{Mail Sending and Transport}
The standard sending and transport protocol is Simple Mail Transfer Protocol (SMTP), first defined in RFC 821 \cite{postel1982rfc} in 1982, and most recently updated in RFC 5321 \cite{klensin2008rfc}. SMTP is an application-layer Internet protocol. In SMTP, a user runs a mail client, or a mail user agent (MUA). This MUA submits mail to a mail server, or a mail submission agent (MSA) on either port 25, 587, or with SSL enabled on 465. The MSA passes the message to a mail transfer agent (MTA), which is often on the same machine as the MSA. The MTA then looks to transfer the message to the recipient's mail exchanger (MX). It does this by looking up the recipient's domain in the Domain Name System (DNS). The MTA then sends the email over the Internet to the MX. The MX then sends the email to the recipient's mail delivery agent (MDA). The MDA will either store the message or continue to forward them to another MDA. Upon arriving at their final MDA, the SMTP portion of email delivery is complete.

\subsubsection{Mail Retrieval, Storage, and Synchronization}
\label{legacyimap}
Once a message has arrived at an MDA, the recipient's MUA must retrieve the message. For this, most MUAs use either POP or IMAP. Post Office Protocol (POP) is a simple protocol that enables downloads and deletions, and is most often used when a single client is interested in retrieving mail from an MDA \cite{reynolds1984post}, \cite{myers1996post}. The Internet Message Access Protocol is a more complex protocol that enables multiple clients to be connected to the same mailbox, and has come into more widespread use in recent years \cite{rfc3501}.

@@

\subsection{Non-metadata End-To-End (E2E) Security}
While the standard email protocols are not secure in and of themselves, there have been numerous attempts over the years to provide secure alternatives. Link encryption (sending over SSL/TLS) has been the most widely adopted, but suggestions for messaging with E2E security have been proposed to varying levels of adoption and with a diverse range of feature inclusion \cite{unger2015sok}. These features include synchronicity, deniability and forward/backward secrecy.

@@

\subsubsection{Messaging Apps}
Some of the most usable and widespread E2E solutions are not for email, but for mobile chat. Apple's iMessage application encrypts messages E2E, although their key management system is unknown and may therefore be invisibly compromised \cite{imessage}. Yet their messaging system is not only usable, it is a widely-adopted example of a smooth transition to an updated protocol without problematic updates to the user experience. Signal and TextSecure, from Open Whisper Systems, provide a similar experience, but avoid iMessage's public key infrastructure dependency by using instead a modified version of Off-the-Record Messaging (OTR), which can use solutions to the socialist millionaires problem in place of relying on a previously known public key \cite{borisov2004off}. The socialist millionaires problem is a zero-knowledge construction where multiple parties would like to test private values for equality without revealing any information to the other party other than the result of the test. Crucially, Signal and TextSecure enable asynchronous messaging by preemptively generating and sending signed key exchange messages \cite{whisper}. They thereby also provide forward secrecy and deniability in an asynchronous setting. The technology used to implement their solution was implemented in the popular messaging application WhatsApp as well \cite{whatsapp}.

@@

These challenges have been addressed more thoroughly in the space of providing transport anonymity, and particularly for Internet routing. Tor, a system for anonymizing Internet traffic, sees about 2.5 million users per day \cite{torusers}. Anonymous remailers use Tor-like forwarding, but have not seen as wide adoption. Pond, a non-email messaging application, provides forward secure asynchronous messaging that does not leak traffic information ``except [to] a global passive attacker." \cite{pond}. There also exist SMTP servers that run as Tor hidden services, essentially bootstrapping their mail anonymity off of Tor's anonymity \cite{onionmail}.

\subsubsection{Mailpile and SMTorP}
While anonymous servers abound, mail clients are generally left to the user's preference. Mailpile, though, is a mail client with a goal of integrating security features into the user experience \cite{mailpile}. It integrates key management into its main contact management system, and includes visual indicators (green locks) to denote secure messages in the style of https in browser URL bars. Furthermore, it contains an experimental plugin (SMTorP) that enables users to opportunistically communicate with other users of the plugin, over Tor \cite{smtorp}. SMTorP is semantically similar to SMTP, except that the receiving client runs an SMTP server on the same machine as the mail client, set up as a Tor hidden service. When used in this fashion, an email is sent to the recipient's .onion address instead of a human-chosen address. The authors note that running an SMTP server on a Tor hidden service might be a security liability, since this opens the client to potentially malicious traffic. Additionally, the alphanumeric string preceding the .onion address is not human-memorable. To address the lack of human-memorability in addresses, the authors consider having users advertise their addresses in the headers of legacy emails that they send, as an invisible upgrade rollout strategy. Clients could then advertise their upgraded status through these headers, invisibly switching to the new protocol when both communicating parties have received the .onion address associated with the legacy email address. The protocol also uses hashcash, a proof-of-work algorithm, to defend against spam (bulk commercial unsolicited email).

Through these tools, the transport mechanism becomes more secure. Yet for all of these, messages are assumed to be transient or remain only on the client device. This is unacceptable behavior to a typical modern user. Modern users expect their emails to be stored on a server and available to them from multiple clients -- from their computers, mobile devices, and on the Web. To ubiquitize secure email, we must develop a storage and synchronization solution that protects messages at rest without compromising the benefits of secure transport protocols.

@@

IMAP is a synchronization protocol that uses a central server to provide mail storage functionality to multiple clients. Popular mail providers such as Gmail, Yahoo! Mail, and Hotmail provide their users with both mail server and client functionality. \project assumes an untrusted network and an untrusted server, but a trusted client. The attacker controls the server, so we do not expect it to guarantee availability. The attacker may both passively analyze any data it is sent for timing or content information, and reply dishonestly to client requests. The IMAP servers of multiple users may collude. A network adversary, either passive or active, may collude with the adversary running the IMAP server. Either of these may themselves be users of \projectnospace, and attempt to gain information from communicating with a target user. All clients of the same account on a mail server have equivalent access privileges. These adversaries may attempt to determine the contents of a message, the communicants, and when a message was received (in order to determine the other two). We make the prior assumption that messages were initially sent to the client in a metadata-secure manner. A client should be able to identify and reject misinformation sent to it by the server.

@@

Our target user is non-technical but experienced with modern email tools. Given a standard mail client, the user can easily send and receive emails, and handle advanced email features like folders and labels. They may need to send files as attachments over emails, and do not wish to receive spam. They access their email from a variety of devices, including their desktop, smartphone, and a web browser. The user must notice no difference between their legacy email experience and their metadata-protected email experience.

\section{Design overview}
To meet these requirements, we present \projectnospace, a client-only extension to the IMAP protocol that provides metadata privacy-preserving storage for multiple \project clients. In \projectnospace, emails received at the client are fully encrypted, including mail headers, indexed, and uploaded to a specialized mailbox in the IMAP server. This hides sender and recipient information from the IMAP server. Encryption is based on a passphrase and generated salt, using scrypt as a key derivation function. A special-use index message is stored in the mailbox as well, and tracks metadata for the client such as the apparent mailbox that a message is in (called a pseudo-mailbox, and visible only to clients, not the server). A special-use lock message is used to allow multiple clients access to the same mailbox, in keeping with IMAP's design principles. Message arrival times are hidden from the server using a combination of scheduling, mixing, fake decoy messages, and techniques drawn from oblivious RAM. \project provides rollback detection for both messages and the index, using version numbers saved client-side and message hashes stored server-side. It is robust to client failure in any state.

@@

Email sizes roughly follow a ``long tail" distribution (where the integral of the graph of distribution of sizes is logarithmic). Sizes below are calculated on a mailbox with a total of 77002 messages, and cumulative number of messages below a particular size are graphed.

@@

The following graphic displays message timestamps per day of the week and hour of the day, in a Github-style ``punchcard" format \cite{punchcard}. In this graphic, day-time combinations that receive more messages are shaded darker, and with larger circles.

@@

We can also see message interarrival times \--- that is, the distribution of the length of time between message arrivals and sends. Since this data was taken from the Gmail ``All Mail" mailbox, it includes both received and sent messages, as the set of all messages synchronized to an IMAP server would. Plotted on a linear scale, the ``long tail" distribution is once again evident; since it has an extremely high slope, we show it also on a logarithmic scale for clarity. The vertical axis shows the total number of messages with inter-message arrival time less than or equal to the corresponding value on the horizontal axis.

@@

A permissible \project solution must be built to integrate with legacy systems. A metadata-secure MUA will also accept legacy SMTP traffic, and it must appear to the user that all messages, once received, are the same. While the user experience must remain consistent, there are multiple possibilities for architecting a storage system consistent with the storage needs. To architect such a backend, there are several possibilities:

@@

The first is unacceptable, given the target user group described in Section~\ref{targetuser}, as this would violate user expectations. Users expect to be able to access their content across all of their devices, and keeping all messages on the local machine only would thereby force them to adjust their usage patterns. To implement the second possibility, we would need to describe a server infrastructure for storing messages and synchronizing across multiple possible clients who have subscribed to updates on messages stored in the system -- which is to say, (a subset of) the functions that the IMAP server is designed to perform. This brings us to our final option, using a modified version of existing IMAP technology, which was described in Section~\ref{legacyimap}.

In the \project protocol, a standard IMAP client library is wrapped with a \project module. An MUA then uses the \project module in lieu of a standard IMAP connection agent. The \project module will perform all necessary functions on behalf of the MUA, including encryption, UID translation, and metadata tracking. The IMAP server is unmodified.
% The \project protocol is 
% The core elements of the \project protocol are 
% \begin{enumerate*}[label=(\itshape\arabic*\upshape)]
% \item an \project module used in lieu of a standard IMAP connection agent in an MUA, and
% \item an index file that is stored as a regular message in a legacy IMAP server. \end{enumerate*}
As further described in later sections, \project ensures that metadata is encrypted along with message contents for all messages to be stored on the IMAP server. It also handles the indexing functions, mapping pseudo-mailboxes to the messages they contain. This is a significant departure from the legacy system, where mailbox names and contents were exposed to the IMAP server. As such, it shifts the burden of tracking, updating, and syncing changes to the message metadata store (the INDEX) from the server to the client. When one client updates the INDEX, the IMAP server will see that a message has changed, and inform its other clients in the standard method.

@@

Note that these changes imply that if one client of an IMAP server is \project-compliant, all clients subscribed to that server must also be \project-compliant. Otherwise, all other clients will see only the encrypted messages as the IMAP server would see them; this is a necessary and desirable feature of the system.

@@

@@

\subsection{Command Translations}
In the IMAP server/client model, the client sends commands to the server, and the server responds based on its state and contents. These commands are the primary form of synchronizing data between the server and its clients.The \project module sits between the client and the legacy IMAP module. It translates arguments, processes data, and returns results that give the client a pseudo-view of the contents of the IMAP server informed by the contents of the metadata INDEX.

We can classify mailboxes by two features: their visibility to the server, and their visibility to the client. When the client's view of the mailbox is used, it is referred to as a pseudo-mailbox. There are four possible mailbox classifications. The classification determines the actions that the module takes in response to each command. A PSEUDO mailbox is not mirrored in the IMAP server; it exists only in the INDEX and is thus only known to clients following the \project protocol. SMTorP is an example of a PSEUDO mailbox. A HYBRID mailbox exists both in the IMAP server and in the INDEX. This is a mailbox such as INBOX, which must also be known the IMAP server as a location for the delivery of incoming messages. A PSEUDO/HIDDEN pseudo-mailbox neither exists in the IMAP server nor is known to the client. It is for use only by the \project module. The FAKE\_MESSAGES pseudo-mailbox used by the scheduler is an example of this type of pseudo-mailbox. An ACTUAL/HIDDEN mailbox is known to the IMAP server but not to the \project client. The CRYPTOBLOBS mailbox, the mailbox in which all encrypted messages are stored on the IMAP server, is an example of this type of mailbox. These classifications are shown in the table below.

@@

In a view of the system enforced by the Python wrapper for an IMAP connection, imaplib, when a SELECT command is passed to the IMAP module, the state of the IMAP server changes to record that a particular client connection has selected a mailbox, and returns a count of the number of messages in that mailbox. When the \project module receives a SELECT command, it saves the requested mailbox to an instance variable. If the requested mailbox is a HYBRID mailbox, it passes the count onto the IMAP server. For both PSEUDO and HYBRID mailboxs, it then calculates the number of messages in the mailbox based on the INDEX, and adds that to either the returned value or 0. If a HIDDEN mailbox is requested, a NO response is returned. Note that the counting functionality is implemented by the EXISTS untagged request as further described in Section~\ref{exists} - the SELECT command additionally elicits the EXISTS server response, whose value is used for the return value of the Python function.

\subsubsection{FETCH} A FETCH or UID FETCH command requests information about a particular message, including either metadata or a subset of message body contents. For both metadata and body, \project first translates the requested UID as described in Section~\ref{uid-translation}. If it is a metadata request, it then forwards the translated command to the IMAP server, and uses the response from the server. Note that metadata here means the metadata the IMAP server uses to keep track of its messages, namely FLAGS, INTERNALDATE, RFC822.SIZE, and ENVELOPE. If it is a body request, \project requests the entire message body (body here includes headers). It then checks that the hash matches, as further described in Section~\ref{rollback}. It then decrypts and checks the authentication of the message, as described in Section~\ref{encryption}. Finally, it returns the appropriate components as specified in the initial request.

@@

\subsubsection{SEARCH} A SEARCH command acts on the selected mailbox. Similarly to SELECT handling as described in Section~\ref{select}, \project returns results from HYBRID and PSEUDO mailboxs as appropriate. The current version of the prototype only responds to SEARCH ALL commands, which list all UIDs in the mailbox in ascending order. By RFC 3501, the IMAP specification, a SEARCH command should handle a wide variety of requests. Many of these would require iterating through each message in the pseudo-mailbox, or searching on encrypted data. While one could imagine protocols for constructing efficient indexes or incorporating homomorphic encryption, that is beyond the scope of this project.

@@

\subsubsection{LIST} The LIST command is used to recursively find all mailboxes in the server. The current prototype allows only top-level, no-children PSEUDO mailboxes. Thus, in translating LIST results, \project inserts PSEUDO mailbox results as if they were to exist at the top level. \project also removes HIDDEN mailboxes from the LIST results before returning.

@@

While all encrypted messages are stored on the IMAP server in the CRYPTOBLOBS mailbox, the client's view should allow for a more complex representation of which messages are stored in which mailboxes. At a minimum, we would like to support an INBOX mailbox for messages that are received at the IMAP server, and an SMTorP or Pond pseudo-mailbox for messages that are received at the client via a metadata-secure transport mechanism. Thus, each message needs an associated UID for its pseudo-mailbox, or the mailbox that the message is in as it appear to the client. But it also has an actual UID, or the UID inside the CRYPTOBLOBS mailbox (the server-view mailbox that stores all messages processed by \projectnospace). This UID is assigned by the IMAP server when \project APPENDs a message to a mailbox on the server. The APPEND command does not, however, return the assigned UID. Therefore we need a separate mechanism for associating a message's pseudo-UID with its actual UID.

To handle this, before appending a message to CRYPTOBLOBS, \project generates a 16-byte random UNIQUE SUBJECT number. This number is placed in the subject line of the appended message, forming an full subject line of the form ``ENCRYPTED <UNIQUE SUBJECT>". Then, when the client requests a message via a specific mailbox/UID pair, \project uses the INDEX to look up the UNIQUE SUBJECT for the pseudo-UID in the appropriate pseudo-mailbox. It then searches for the message in CRYPTOBLOBS with UNIQUE SUBJECT in its subject line, which returns the unique actual UID of the message in the CRYPTOBLOBS mailbox.

@@

\subsection{Processing Messages Received at IMAP Server}
\label{processing}

@@

Because of this delete/upload lock acquisition mechanism, the IMAP server may sometimes be in a state where the LOCK message does not exist but no client holds the lock, such as if a client crashes or closes down while holding the lock. Optimally, this would not happen, and the lock would be released before closing. Unfortunately, the existing library allows its client to close its socket directly, rather than through a specified method at the library level, so we cannot rely on a client to release a lock before quitting. Thus, \project includes a mechanism for detecting this inconsistent state and recovering from it by acquiring a lock even if the LOCK message does not exist in the CRYPTOBLOBS mailbox.

>>>>

\begin{abstract}
Although communication metadata is known to be intrusive, most end-to-end encryption solutions focus only on hiding the contents of messages. Yet who a user is talking to and when is information as sensitive as the contents of the message. While end-to-end metadata security solutions exist for message transport, existing solutions lack the multi-device synchronization that IMAP provides. Without providing full feature parity with legacy email technologies, a metadata-secure protocol cannot hope to achieve ubiquity. To that end, we present \projectnospace, a client-side extension of the IMAP protocol held to rigorous standards of feature parity without sacrificing on security. With \projectnospace, existing IMAP clients can be invisibly upgraded to provide metadata security for email storage and synchronization, enabling the rollout of end-to-end metadata secure messaging.
\end{abstract}

@@

Today, most email is hosted by cloud providers. Unfortunately for users, this means that information about who they're talking to, when they're talking to those people, and what they're saying to each other is readily available to any interested company, government, or advertiser. Given that a modern threat model must account for the mail provider as an adversary, then only an end-to-end (E2E) solution will provide adequate security. Furthermore, even encrypting the contents of the email will protect only a small portion of the full breadth of sensitive information: the metadata of who is communicating and when is equally important to protect. Fortunately, recent projects have emerged to add protections for metadata into the communication security toolkit. Yet these projects lack a key feature to achieve full feature parity with insecure messaging solutions: support for multi-client synchronization and storage across all of a user's devices.

@@

While current technologies do not provide features that satisfy our stated goals, existing protocols and applications form the ecosystem in which \project resides.

@@

To send an email, the sender's machine sends a request to an SMTP server acting on its behalf, which relays the message through other SMTP servers until it reaches the recipient's destination server. At that point, the recipient's machine retrieves the email from the server using either POP or IMAP.

\subsubsection{Mail Sending and Transport}
The standard sending and transport protocol is Simple Mail Transfer Protocol (SMTP), first defined in RFC 821 \cite{postel1982rfc} in 1982, and most recently updated in RFC 5321 \cite{klensin2008rfc}. SMTP is an application-layer Internet protocol. In SMTP, a user runs a mail client, called a mail user agent (MUA). This MUA submits mail to a mail server (the mail submission agent, MSA) on either port 25, 587, or with SSL enabled on 465. The MSA passes the message to a mail transfer agent (MTA), which is often on the same machine as the MSA. The MTA then looks to transfer the message to the recipient's mail exchanger (MX). It does this by looking up the recipient's domain in the Domain Name System (DNS). The MTA then sends the email over the Internet to the MX. The MX then sends the email to the recipient's mail delivery agent (MDA). The MDA will either store the message or continue to forward it to another MDA. Upon arriving at its final MDA, the SMTP portion of email delivery is complete.

\subsubsection{Mail Retrieval, Storage, and Synchronization}
\label{legacyimap}
Once a message has arrived at an MDA, the recipient's MUA must retrieve the message. For this, most MUAs use either POP or IMAP. Post Office Protocol (POP) is a simple protocol that enables downloads and deletions, and is most often used when a single client is interested in retrieving mail from an MDA \cite{reynolds1984post} \cite{myers1996post}. The Internet Message Access Protocol (IMAP) is a more complex protocol that enables multiple clients to be connected to and modifying the same mailbox, and has come into more widespread use in recent years \cite{rfc3501}.

@@

\subsection{Non-metadata End-To-End (E2E) Security}
While the standard email protocols are not secure in and of themselves, there have been numerous attempts over the years to provide secure alternatives. Link encryption (sending over SSL/TLS) has been the most widely adopted, but suggestions for messaging with E2E security have been proposed to varying levels of adoption and with a diverse range of feature inclusion \cite{unger2015sok}. These features include synchronicity, deniability, and forward/backward secrecy.

@@

\subsubsection{Messaging Apps}
Some of the most usable and widespread E2E solutions are not for email, but for mobile chat. Apple's iMessage application encrypts messages E2E, although their key management system is unknown and may therefore be invisibly compromised \cite{imessage}. Their messaging system is not only usable: it is a widely-adopted example of a smooth transition to an updated protocol without problematic updates to the user experience. Signal and TextSecure, from Open Whisper Systems, provide a similar experience, but avoid iMessage's public key infrastructure dependency by using instead a modified version of Off-the-Record Messaging (OTR), which can use solutions to the socialist millionaires problem in place of relying on a previously known public key \cite{borisov2004off}. The socialist millionaires problem is a zero-knowledge construction where multiple parties would like to test private values for equality without revealing any information to the other party other than the result of the test. Crucially, Signal and TextSecure enable asynchronous messaging by preemptively generating and sending signed key exchange messages \cite{whisper}. They thereby also provide forward secrecy and deniability in an asynchronous setting. The technology used to implement their solution was implemented in the popular messaging application WhatsApp as well \cite{whatsapp}.

@@

These challenges have been addressed more thoroughly in the space of providing transport anonymity, and particularly for Internet routing. Tor, a system for anonymizing Internet traffic, sees about 2.5 million users per day \cite{torusers}. As these figures have taught us, Tor users demonstrate a desire to hide the metadata of their online activities. Anonymous remailers use Tor-like forwarding, but have not seen as wide adoption. Pond, a non-email messaging application, provides forward secure asynchronous messaging that does not leak traffic information ``except [to] a global passive attacker.'' \cite{pond}. There also exist SMTP servers that run as Tor hidden services, essentially bootstrapping their mail anonymity off of Tor's anonymity \cite{onionmail}.

\subsubsection{Mailpile and SMTorP}
While anonymous servers abound, mail clients are generally left to the user's preference. Mailpile, in contrast, is a mail client designed with a goal of integrating security features into the user experience \cite{mailpile}. It integrates key management into its main contact management system, and includes visual indicators (green locks) to denote secure messages in the style of https in browser URL bars. Furthermore, Mailpile contains an experimental plugin (SMTorP) that enables users to opportunistically communicate with other users of the plugin in a metadata-secure manner, by sending messages over Tor \cite{smtorp}. SMTorP is semantically similar to SMTP, except that the receiving client runs an SMTP server on the same machine as the mail client, set up as a Tor hidden service. When used in this fashion, an email is sent to the recipient's .onion address (the address of the hidden service) instead of a human-chosen address. The Mailpile authors note that running an SMTP server on a Tor hidden service might be a security liability, since this opens the client to potentially malicious traffic. Additionally, the alphanumeric string preceding the .onion address is not human-memorable. To address the lack of human-memorability in addresses, the authors consider having users advertise their addresses in the headers of legacy emails that they send, as an invisible upgrade rollout strategy. Clients could then advertise their upgraded status through these headers, invisibly switching to the new protocol when both communicating parties have received the .onion address associated with the legacy email address. The protocol also uses hashcash, a proof-of-work algorithm, to defend against spam (bulk commercial unsolicited email).

Through these tools, the transport mechanism becomes more secure. Yet for all of these, messages are assumed to be transient or remain only on the client device. This is unacceptable behavior to a typical modern user. Modern users expect their emails to be stored on a server and available to them from multiple clients -- from their computers, mobile devices, and on the Web. We could perhaps imagine a situation without an IMAP-like protocol, where the machine that receives messages is also a trusted webmail server and therefore need not perform any synchronization. Although webmail is popular, such a solution is still a feature rollback, and is therefore unacceptable ideologically given our stated aims. To ubiquitize secure email, we must develop a storage and synchronization solution that protects messages at rest without compromising the benefits of secure transport protocols.

@@

Given the modern state of pervasive surveillance, it is appropriate for our threat model to err on the side of paranoia. The IMAP protocol uses a central server to provide mail storage functionality to multiple clients. Popular mail providers such as Gmail, Yahoo! Mail, and Hotmail provide their users not only with servers, but with client functionality as well. \project assumes an untrusted network and an untrusted server, but a trusted client. The attacker controls the server, so we do not expect it to guarantee availability. The attacker may both passively analyze any data it is sent for timing or content information, and reply dishonestly to client requests. The IMAP servers of multiple users may collude. A network adversary, either passive or active, may collude with the adversary running the IMAP server. Either of these may themselves be users of \projectnospace, and attempt to gain information from communicating with a target user. All clients of the same account on a mail server have equivalent access privileges. These adversaries may attempt to determine the contents of a message, the communicants, and when a message was received (in order to determine the other two). We make the prior assumption that messages were initially sent to the client in a metadata-secure manner. A client should be able to identify and reject misinformation sent to it by the server.

@@

To achieve ubiquitous adoption, we must not place high technical expectations on our target user. Our target user is non-technical but experienced with modern email tools. Given a standard mail client, the user can easily send and receive emails, and handle advanced email features like folders and labels. They may need to send files as attachments over emails, and do not wish to receive spam. They access their email from a variety of devices, including their desktop, smartphone, and a web browser. The user must notice no difference between their legacy email experience and their metadata-protected email experience.

\section{Design overview}
To meet all aforementioned requirements, we present \projectnospace, a client-only extension to the IMAP protocol that provides metadata privacy-preserving storage for multiple \project clients. In \projectnospace, emails received at the client are fully encrypted (including mail headers), indexed, and uploaded to a specialized mailbox in the IMAP server. This hides sender and recipient information from the IMAP server. Encryption is based on a passphrase and generated salt, using scrypt as a key derivation function. A special-use ``INDEX'' message is stored in the mailbox as well, and tracks metadata for the client such as the apparent mailbox that a message is in (called a pseudo-mailbox, and visible only to clients, not the server). A special-use ``LOCK'' message is used to allow multiple clients access to the same mailbox, in keeping with IMAP's design principles. Message arrival times are hidden from the server using a combination of scheduling, mixing, fake decoy messages, and techniques drawn from oblivious RAM. \project provides rollback detection for both messages and the index, using version numbers saved client-side and message hashes stored server-side. It is robust to client failure in any state.

@@

Email sizes roughly follow a ``long tail'' distribution (where the integral of the graph of distribution of sizes is logarithmic). Sizes below are calculated on a mailbox with a total of over 77,000 messages, and cumulative number of messages below a particular size are graphed.

@@

The following graphic displays message timestamps per day of the week and hour of the day, in a Github-style ``punchcard'' format \cite{punchcard}. In this graphic, day-time combinations that receive more messages are shaded darker, and with larger circles.

@@

We can also find message interarrival times \---- that is, the distribution of the length of time between message arrivals and sends. Since this data was taken from the Gmail ``All Mail'' mailbox, it includes both received and sent messages, as the set of all messages synchronized to an IMAP server would. Plotted on a linear scale, the ``long tail'' distribution is once again evident; since it has an extremely high initial slope, we show it also on a logarithmic scale for clarity. The vertical axis shows the total number of messages with inter-message arrival time less than or equal to the corresponding value on the horizontal axis.

@@

A permissible \project solution must be built to integrate with legacy systems. A metadata-secure client will also accept legacy SMTP traffic, and it must appear to the user that all messages, once received, are functionally the same. While the user experience must remain consistent, there are multiple possibilities for architecting a storage system consistent with the storage needs. To architect such a backend, there are several possibilities:

@@

The first is unacceptable, given the target user group described in Section~\ref{targetuser}, as this would violate user expectations. Users expect to be able to access their content across all of their devices, and keeping all messages on the local machine only would thereby force them to adjust their usage patterns. To implement the second possibility, we would need to describe a server infrastructure for storing messages and synchronizing across multiple possible clients who have subscribed to updates on messages stored in the system -- which is to say, (a subset of) the functions that the IMAP server is designed to perform. This brings us to our final option: using a modified version of existing IMAP technology, as we have described.

In the \project protocol, a standard IMAP client library is wrapped with an \project module. A client then uses the \project module in lieu of a standard IMAP connection agent. The \project module will perform all necessary functions on behalf of the client, including encryption, message ID translation, and metadata tracking. The IMAP server is unmodified.
% The \project protocol is 
% The core elements of the \project protocol are 
% \begin{enumerate*}[label=(\itshape\arabic*\upshape)]
% \item an \project module used in lieu of a standard IMAP connection agent in an MUA, and
% \item an index file that is stored as a regular message in a legacy IMAP server. \end{enumerate*}
As further described in later sections, \project ensures that metadata is protected along with message contents for all messages to be stored on the IMAP server. It also handles the indexing functions, mapping pseudo-mailboxes to the messages they contain. This is a significant departure from the legacy system, where mailbox names and contents were exposed to the IMAP server. As such, it shifts the burden of tracking, updating, and syncing changes to the message metadata store (the INDEX) from the server to the client. When one client updates the INDEX, the IMAP server will receive the modified INDEX, making these changes visible to all other clients.

@@

Note that these changes imply that if one client of an IMAP server is \projectnospace-compliant, all clients subscribed to that server must also be \projectnospace-compliant. Otherwise, all other clients will see only the encrypted messages as the IMAP server would see them; this is a necessary and desirable feature of the system.

@@

Here we present the details of the system as it is currently implemented. While it currently implements only a subset of RFC 3501-specified functionality, it satisfies all requirements of Mailpile, its functioning client for prototyping purposes.

@@

\subsection{Command Translations}
In the IMAP server/client model, the client sends commands to the server, and the server responds based on its state and contents. These commands are the primary form of synchronizing data between the server and its clients. The \project module sits between the client and the legacy IMAP module. It translates arguments, processes data, and returns results that give the client a pseudo-view of the contents of the IMAP server informed by the contents of the metadata INDEX.

We can classify mailboxes by two features: their visibility to the server, and their visibility to the client. When the client's view of the mailbox is used, it is referred to as a pseudo-mailbox. There are four possible mailbox classifications, as illustrated in the table below. The classification determines the actions that the module takes in response to each command. A PSEUDO mailbox is not mirrored in the IMAP server; it exists only in the INDEX and is thus only known to clients following the \project protocol. SMTorP is an example of a PSEUDO mailbox. A HYBRID mailbox exists both in the IMAP server and in the INDEX. This is a mailbox such as INBOX, which must also be known to the IMAP server as a location for the delivery of incoming messages. A PSEUDO/HIDDEN pseudo-mailbox neither exists in the IMAP server nor is known to the client. It is only for the internal use of the \project module. The FAKE\_MESSAGES pseudo-mailbox used by the scheduler is an example of this type of pseudo-mailbox. An ACTUAL/HIDDEN mailbox is known to the IMAP server but not to the \project client. The CRYPTOBLOBS mailbox, the mailbox in which all encrypted messages are stored on the IMAP server, is an example of this type of mailbox. These classifications are shown in this table.

@@

In a view of the system enforced by the Python wrapper for an IMAP connection, imaplib, when a SELECT command is passed to the IMAP module, the state of the IMAP server changes to record that a particular client connection has selected a mailbox, and returns a count of the number of messages in that mailbox. When the \project module receives a SELECT command, it saves the requested mailbox to an instance variable. If the requested mailbox is a HYBRID mailbox, it passes the count onto the IMAP server. For both PSEUDO and HYBRID mailboxes, \project then calculates the number of messages in the mailbox based on the INDEX, and adds that to either the returned value or 0. If a HIDDEN mailbox is requested, a NO response is returned. Note that the counting functionality is implemented by the EXISTS untagged request as further described in Section~\ref{exists} \---- the SELECT command additionally elicits the EXISTS server response, whose value is used for the return value of the Python function.

\subsubsection{FETCH} A FETCH or UID FETCH command requests information about a particular message, including either metadata or a subset of message body contents. For both metadata and body, \project first translates the requested permanent message ID (UID) as described in Section~\ref{uid-translation}. If it is a metadata request, it then forwards the translated command to the IMAP server, and forwards the response from the server. Note that metadata here means the metadata the IMAP server uses to keep track of its messages, namely FLAGS, INTERNALDATE, RFC822.SIZE, and ENVELOPE. If it is a body request, \project requests the entire message body (where body also includes headers). \project then checks that the hash matches, as further described in Section~\ref{rollback}. It then decrypts and checks the authentication of the message, as described in Section~\ref{encryption}. Finally, it returns the appropriate components as specified in the initial request.

@@

\subsubsection{SEARCH} A SEARCH command acts on the selected mailbox. Similarly to SELECT handling as described in Section~\ref{select}, \project returns results from HYBRID and PSEUDO mailboxes as appropriate. The current version of the prototype only responds to SEARCH ALL commands, which list all UIDs in the mailbox in ascending order. By RFC 3501, the IMAP specification, a SEARCH command should handle a wide variety of requests. Many of these would require iterating through each message in the pseudo-mailbox, or searching on encrypted data. While one could imagine protocols for constructing efficient indexes or incorporating homomorphic encryption, such abilities are beyond the scope of this project.

@@

\subsubsection{LIST} The LIST command is used to recursively find all mailboxes in the server. The current prototype allows only top-level, no-children PSEUDO mailboxes. Thus, in translating LIST results, \project inserts PSEUDO mailbox results as if existed at the top level. \project also removes HIDDEN mailboxes from the LIST results before returning.

@@

While all encrypted messages are stored on the IMAP server in the CRYPTOBLOBS mailbox, the client's view should allow for a more complex representation of which messages are stored in which mailboxes. At a minimum, we would like to support an INBOX mailbox for messages that are received at the IMAP server, and an SMTorP or Pond pseudo-mailbox for messages that are received at the client via a metadata-secure transport mechanism. Thus, each message needs an associated UID for its pseudo-mailbox, or the mailbox that the message is in as it appears to the client. But it also has an actual UID, or the UID inside the CRYPTOBLOBS mailbox (the server-view mailbox that stores all messages processed by \projectnospace). This UID is assigned by the IMAP server when \project APPENDs a message to a mailbox on the server. The APPEND command does not, however, return the assigned UID. Therefore we need a separate mechanism for associating a message's pseudo-UID with its actual UID.

To handle this, before appending a message to CRYPTOBLOBS, \project generates a 16-byte random UNIQUE SUBJECT number. This number is placed in the subject line of the appended message, forming a full subject line of the form ``ENCRYPTED <UNIQUE SUBJECT>''. Then, when the client requests a message via a specific mailbox/UID pair, \project uses the INDEX to look up the UNIQUE SUBJECT for the pseudo-UID in the appropriate pseudo-mailbox. It then searches for the message in CRYPTOBLOBS with UNIQUE SUBJECT in its subject line, which returns the unique actual UID of the message in the CRYPTOBLOBS mailbox.

@@

\subsection{Processing Messages Received at the IMAP Server}
\label{processing}

@@

Because of this delete/upload lock acquisition mechanism, the IMAP server may sometimes be in a state where the LOCK message does not exist but no client holds the lock, such as if a client crashes or closes down while holding the lock. Optimally, this would not happen, and the lock would be released before closing. Unfortunately, the existing library (imaplib) allows its client to close its socket directly, rather than through a specified method at the library level, so we cannot rely on a client to release a lock before quitting. Thus, \project includes a mechanism for detecting this inconsistent state and recovering from it by acquiring a lock even if the LOCK message does not exist in the CRYPTOBLOBS mailbox.
